Towards open collaboration in biomedical and health research: An open collaborative scoping review

### 2025-03-18 doc/paper.qmd
Write results section - laboratory-level practices (#118)
Added implementation of open collaboration practices at laboratory level

---------

Co-authored-by: Luke W. Johnston <lwjohnst@gmail.com>

In @Turoman2022 the authors describe in detail how the have set up a
laboratory whose members collaborate in an open way from study design to
dissemination. Here we highlight the role of group meetings as
opportunities to propose ideas and have an initial study design
proposed. These group meetings are usually started with a section of
interesting mistakes, which allows to share the issues, troubleshoot and
make the laboratory members comfortable with the fact that errors are
unavoidable in science. The feedback from these meetings might lead to
updates to a lab manual, which is a great tool for new members to start
engaging with the open practices of the lab. This coupled with a support
system by senior members of the laboratory allows, in theory, to ease
the learning curve of new members to adapt to open practices and
collaboration.

In the paper, authors describe their implementation of open practices in
each step of the scientific production. Here we focus on the aspects in
pre-registration, data analysis and dissemination that improve
collaboration between members of the laboratory and of the same field.
For pre-registration they combine slack channels and Trello to
communicate between members and OSF for data deposition and uploading
pre-registered studies. They put special emphasis on folder structure,
which is consistent for all projects, and assigning the authorship in
this stage. For data analysis, the authors highlight the figure of the
data partner, a member of the laboratory not originally included in the
analyses that runs the analyses described in the pre-registered study
independently. If the results mismatch with the original ones meetings
with the lead analyst or all collaborators are organized to assess what
the issue might be. Additionally, this allows for insights that might
result in deviations of the original pre-registered study, which should
be included and noted in the final manuscript. Importantly, the role of
the data partner is described in the final manuscript. Finally, in the
dissemination step, they highlight that they might incorporate feedback
from presenting their analyses in conferences, which are often included
as additional non-preregistered analyses.

Finally, in the paper they also highlight relevant open practices that
might not enhance collaboration, but that we consider interesting
reviewing too. First, during pre-registration, they allow for papers to
contain defined confirmatory and exploratory sections to avoid the usual
rigidness of pre-registered studies. Second, in exploratory analyses
they included soundness checks whenever possible. Third, the perform
sequential hypothesis testing to assess within the limits of a maximum
sample size which hypothesis fits the data better without indulging in
p-hacking. Finally, they collaborate with schools and with universities
to disseminate their results.

>>>>

In @Turoman2022, the authors describe in detail how they set up a
laboratory where members collaborate in an open way, from study design
to dissemination. Here we highlight the role of group meetings as
opportunities to propose ideas and discuss an initial study design.
These group meetings usually start with some time to reflect on and
discuss interesting mistakes, which allows members to share the issues
and potential troubleshoot the problems. This helps make the laboratory
members feel more comfortable with experiencing and managing errors, as
they are unavoidable in science. The feedback from these meetings might
lead to updating a lab manual, which is a great tool for new members to
start engaging with the open practices of the lab. This, coupled with a
support system by senior laboratory members allows, at least in theory,
a smoother on-boarding experience for new members to adapt to open
practices and collaboration.

In they paper they describe how they implement open practices in each
step of the scientific production. We will focus on the aspects on
pre-registration, data analysis, and dissemination that improve
collaboration between members of the laboratory and of the same field.
For pre-registration they combine Slack channels and Trello to
communicate within the group and use OSF to deposit data and upload
pre-registered studies. They put special emphasis on folder structure,
which is consistent for all projects, and assigning the authorship at
this stage. For data analysis, they incorporate a "data partner" role,
who is a member of the laboratory not originally included in the study.
This data partner independently runs the analyses described in the
pre-registered study. If there is a mismatch in results with the
original results, meetings with the lead analyst or all collaborators
are organized to assess what the issue might be. This allows for
insights that may result in deviations in the original pre-registered
study, which should be included and noted in the final manuscript. The
role of the data partner is, importantly, included and described in the
final manuscript. In the final dissemination step, they highlight that
they may incorporate feedback given after presenting their analyses in
conferences, which are often included as additional non-preregistered
analyses.

The group also highlights that relevant open practices may not always
enhance collaboration, but have other benefits that we want to also
highlight. First, during pre-registration, they allow for papers to
contain defined confirmatory and exploratory sections to avoid the usual
rigidness of pre-registered studies. Second, in exploratory analyses
they included soundness checks whenever possible. Third, they perform
sequential hypothesis testing to assess within the limits of a maximum
sample size which hypothesis fits the data better without indulging in
p-hacking. Finally, they collaborate with schools and with universities
to disseminate their results.
